---
title: "**Evaluation of Labor Policy**"
subtitle: "Assignment 2 - Difference-in-Differences"
author: "Henri Bui, Grace Yan, Samy Ghnaïa"
date: \today
output: html_document
# html_document: 
#   theme: sandstone
#   highlight: tango
# pdf_document: default
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(dplyr)
library(tidyr)
library(knitr)
library(fixest)
library(ggplot2)
library(gt)

# free the working memory
rm(list = ls())

# set working directory 
setwd(here::here())

# set seed 
set.seed(1234)

# import the data in an object called 'data'
data <- read.csv("data2.csv")
```

<br>

## 1. Descriptive evidence

**a. Create a categorical variable `treatment_group` for the cohorts in the data**

```{r new variable treatment_group, include=TRUE}

data <- data %>%
  mutate(treatment_group = case_when(
    agegrp >= 45 ~ 1,
    (agegrp <= 44 & agegrp >= 35) ~ 2,
    (agegrp <= 34 & agegrp >= 25) ~ 3,
    (agegrp <= 24 & agegrp >= 20) ~ 4
  ))
```

The `treatment group` variable now takes values: <br>
  - 1 for cohorts aged 45 or older,<br>
  - 2 for cohorts aged or between 35-44 years old,<br>
  - 3 for cohorts aged or between 25-34 years old,<br>
  - 4 for cohorts aged or between 20-24 years old.<br>

<br>

**b. Produce bar plots of mean `education` and mean `primary` by `treatment_group`**

```{r bar plots, include=TRUE}

data <- data %>%
  mutate(treatment_label = case_when(
    treatment_group == 1 ~ "Pre-treatment\n(45+)",
    treatment_group == 2 ~ "Early treated\n(35-44)",
    treatment_group == 3 ~ "Late treated\n(25-34)",
    treatment_group == 4 ~ "Post-treatment\n(20-24)"
  ))

# calculate means by treatment group
means <- data %>%
  group_by(treatment_group, treatment_label) %>%
  summarise(
    mean_education = mean(education, na.rm = TRUE),
    mean_primary = mean(primary, na.rm = TRUE)
  )

# bar plot for education
education_plot <- ggplot(means, aes(x = reorder(treatment_label, treatment_group), 
                  y = mean_education)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  labs(
    title = "Average Years of Schooling Attained",
    x = "Treatment group",
    y = "Mean years of education"
  ) +
  theme_minimal()

# bar plot for primary schooling
primary_schooling_plot <- ggplot(means, aes(x = reorder(treatment_label, treatment_group), 
                  y = mean_primary)) +
  geom_bar(stat = "identity", fill = "coral") +
  labs(
    title = "Share of Adults with any Primary Schooling",
    x = "Treatment group",
    y = "Share with any primary school"
  ) +
  theme_minimal()

education_plot
primary_schooling_plot

```

On both the ´Average Years of Schooling Attained´ and ´Share of Adults with any Primary Schooling´graphs we observe a similar trend: the post-treatment group displays higher years of schooling attained and a higher share of adults with any primary schooling. Additionally, from the pre-treatment group until the post-treatment group, we see a consistent increase in both the average years of schooling attained and the share of adults with any primary schooling. When it comes to this gradual increase, we observe a significant jump on both graphs between the pre-treatment group and the early-treated group.This is consistent with the 1967 lifting of restrictions having had a positive effect on educational attainment.

<br>

## 2. Validity checks

**a. Means tables (for `education` and `primary`) using pre-treatment and late-treated cohorts**

```{r means tables, include=TRUE}
# create subset
data_subset <- data %>%
  filter(treatment_group == 1 | treatment_group == 3) %>%
  mutate(
    cohort_label = if_else(treatment_group == 1, "Pretreatment", "Late treated"),
    wenela_label = if_else(n_wenela > 0, "Yes", "No")
  )

means <- data_subset %>%
  group_by(cohort_label, wenela_label) %>%
  summarise(mean_val = mean(education, na.rm = TRUE), .groups = "drop") %>%
  pivot_wider(names_from = wenela_label, values_from = mean_val) %>%
  mutate(Diff = Yes - No)

diff_row <- tibble(
  cohort_label = "Diff.",
  No   = means$No[means$cohort_label == "Late treated"] - means$No[means$cohort_label == "Pretreatment"],
  Yes  = means$Yes[means$cohort_label == "Late treated"] - means$Yes[means$cohort_label == "Pretreatment"],
  Diff = NA_real_
)

p_edu_pre  <- t.test(education ~ wenela_label, data = filter(data_subset, cohort_label == "Pretreatment"))$p.value
p_edu_late <- t.test(education ~ wenela_label, data = filter(data_subset, cohort_label == "Late treated"))$p.value

result <- means %>%
  mutate(Diff = Yes - No, p_value = c(p_edu_pre, p_edu_late), Statistical_Significance = c("***", "***")) %>%
  bind_rows(diff_row)

cat("Table 1: Average Years of Schooling\n")
result

means <- data_subset %>%
  group_by(cohort_label, wenela_label) %>%
  summarise(mean_val = mean(primary, na.rm = TRUE), .groups = "drop") %>%
  tidyr::pivot_wider(names_from = wenela_label, values_from = mean_val) %>%
  mutate(Diff = Yes - No)

diff_row <- tibble(
  cohort_label = "Diff.",
  No   = means$No[means$cohort_label == "Late treated"] - means$No[means$cohort_label == "Pretreatment"],
  Yes  = means$Yes[means$cohort_label == "Late treated"] - means$Yes[means$cohort_label == "Pretreatment"],
  Diff = NA_real_
)

p_pri_pre  <- t.test(primary ~ wenela_label, data = filter(data_subset, cohort_label == "Pretreatment"))$p.value
p_pri_late <- t.test(primary ~ wenela_label, data = filter(data_subset, cohort_label == "Late treated"))$p.value

result_primary <- means %>%
  mutate(Diff = Yes - No, p_value = c(p_pri_pre, p_pri_late), Statistical_Significance = c("***", "**")) %>%
  bind_rows(diff_row)

cat("Table 2: Share of Adults by any Primary Schooling\n")
result_primary
```



<br>

**b. Compare pre-treatment vs late-treated cohorts in districts with Wenela stations**

LT is much higher than LT, meaning that the educational attainment of Malawi individuals is higher in the late treated group. We can conclude that the abolishment of quotas in 1967 (migration) has a positive correlation with the educational attainment of Malawi citizens. However, we cannot establish causality between the aforementioned two elements as there may have been other factors at play such that also influence education attainment. It could be that with time, people decide to take their education more seriously or that with time, Malawi's industry shifted towards sectors needing human capital. Therefore, these results cannot be interpreted as causal evidence of migration. 



Written by Claude:

In districts with Wenela stations, late treated cohorts have higher educational attainment than pretreatment cohorts. This is consistent with the income from migration having a positive effect on schooling. However this cannot be interpreted as causal evidence because we are only looking at one group over time — we have no counterfactual. We don't know what would have happened to the late treated cohort's education in the absence of the migration shock. Other time-varying factors could explain the difference, such as general economic development in Malawi over time or improvements in school access.

<br>

**c. Compare districts with vs without Wenela stations among late-treated cohorts**

Comparing districts with and without Wendela stations for the late treated cohort, we observe a higher level of educational attainment for districts with Wendela stations than those without. Here again, we cannot conclude a causal evidence of the effect of migration as there may be structural differences between districts with and without Wenela stations. We can observe a positive correlation but we cannot conclude causality.

Written by Claude:

Among late treated cohorts, districts with Wenela stations have higher educational attainment than those without. Again this cannot be interpreted causally because Wenela and non-Wenela districts may differ in many ways beyond migration access — for example Wenela districts tend to be at higher altitude, have lower population density, and fewer agricultural estates. These structural differences could independently affect education, so we cannot attribute the gap to migration alone.

<br>

**d. Main threats to identification**

The main threat to identification is the parallel trend assumption. We cannot be sure that both districts with and without Wenela treatment would have followed the same trend before the treatment. The districts may be structurally different. If districts with and without Wenela stations did not follow the same trends before the treatment then our DiD estimation (used to establish causality) would include some bias.

Written by Claude:

The main threat is the parallel trends assumption — for DiD to give a causal estimate, Wenela and non-Wenela districts must have followed the same education trend in the absence of the migration shock. If Wenela districts were already on a steeper upward trajectory in education before 1967 for unrelated reasons, our estimate would be biased upward. A concrete example: if Wenela districts happened to receive more missionary schools or government investment in education around the same time as the migration shocks, we could not separate the effect of migration income from the effect of better school supply.

<br>

## 3. Causal effect estimation

**a. Estimate the simple DiD model (pre-treatment and late-treated cohorts only)**

```{r simple did model, echo=TRUE}


# create the dummy variables
data_did <- data_subset %>%
  mutate(
    late_treated = if_else(treatment_group == 3, 1, 0),
    any_wenela   = if_else(n_wenela > 0, 1, 0)
  )

# estimate the DiD model for education
did_education <- lm(education ~ late_treated * any_wenela, data = data_did)
summary(did_education)

# estimate the DiD model for primary
did_primary <- lm(primary ~ late_treated * any_wenela, data = data_did)
summary(did_primary)

```

<br>

**b. What is the interpretation of each coefficient? What can you conclude? **

Interpretation of coefficients: <br><br>
<br>$$\alpha_0$$: This is the intercept and represents the baseline, a.k.a. the level of the dependent $$Y_{cdg}$$ when all other coefficients are equal to 0.
<br>$$\alpha_1$$: Represents the effect of the independent variable, treated group in districts with Wenela stations on the dependent variable, $$Y_{cdg}$$
<br>$$\alpha_2$$: Represents the effect of the dummy variable $$LateTreated_c$$ 
<br>$$\alpha_3$$: 

<br>

**c. External validity**

Your answer.

<br>

## 4. AI Question

```{r chatGPT_code, echo=TRUE}

# --------------------------------------------------
# 1. Load libraries
# --------------------------------------------------
library(tidyverse)
library(fixest)      # for FE regressions
library(modelsummary)

# --------------------------------------------------
# 2. Load data
# --------------------------------------------------
data <- read_csv("data2.csv")

# --------------------------------------------------
# 3. Create cohort treatment groups
# agegrp = 5-year age group in 1998
# --------------------------------------------------

data <- data %>%
  mutate(
    early_treat = ifelse(agegrp %in% c(35,40), 1, 0),
    late_treat  = ifelse(agegrp %in% c(25,30), 1, 0),
    post_treat  = ifelse(agegrp %in% c(20), 1, 0)
  )

# --------------------------------------------------
# 4. Create interactions with Wenela stations
# --------------------------------------------------

data <- data %>%
  mutate(
    early_wenela = early_treat * n_wenela,
    late_wenela  = late_treat  * n_wenela,
    post_wenela  = post_treat  * n_wenela
  )

# --------------------------------------------------
# 5. Create time trend for interactions
# --------------------------------------------------

data <- data %>%
  mutate(
    trend = agegrp
  )

# --------------------------------------------------
# 6. TABLE 5 – Column (1)
# Years of schooling
# --------------------------------------------------

m1 <- feols(
  education ~ early_wenela + late_wenela + post_wenela +
    female + logpopdens31 + frac_literate_5_to18_1945 +
    central + southern |
    district + agegrp,
  cluster = ~district,
  data = data
)

# --------------------------------------------------
# 7. TABLE 5 – Column (2)
# With trend interactions
# --------------------------------------------------

m2 <- feols(
  education ~ early_wenela + late_wenela + post_wenela +
    female + logpopdens31 + frac_literate_5_to18_1945 +
    central + southern +
    trend:logpopdens31 +
    trend:frac_literate_5_to18_1945 +
    trend:central +
    trend:southern |
    district + agegrp,
  cluster = ~district,
  data = data
)

# --------------------------------------------------
# 8. TABLE 5 – Column (3)
# Share with any primary school
# --------------------------------------------------

m3 <- feols(
  primary ~ early_wenela + late_wenela + post_wenela +
    female + logpopdens31 + frac_literate_5_to18_1945 +
    central + southern |
    district + agegrp,
  cluster = ~district,
  data = data
)

# --------------------------------------------------
# 9. TABLE 5 – Column (4)
# With trend interactions
# --------------------------------------------------

m4 <- feols(
  primary ~ early_wenela + late_wenela + post_wenela +
    female + logpopdens31 + frac_literate_5_to18_1945 +
    central + southern +
    trend:logpopdens31 +
    trend:frac_literate_5_to18_1945 +
    trend:central +
    trend:southern |
    district + agegrp,
  cluster = ~district,
  data = data
)

# --------------------------------------------------
# 10. Display table
# --------------------------------------------------

modelsummary(
  list(
    "Years schooling (1)" = m1,
    "Years schooling (2)" = m2,
    "Primary share (3)"   = m3,
    "Primary share (4)"   = m4
  ),
  stars = TRUE
)

```

**a. Identify issues**

### We observe 5 issues/weaknesses concerning the ChatGPT code: <br>

**<br>1. The age groups differ from those in the paper.**

Dinkelman and Mariotti define four age categories:

<br>  - Post treatment cohort: 20-24,
<br>  - Late treated cohort: 25-35,
<br>  - Early treated cohort: 35-44,
<br>  - Pretreatment cohort: 45-64.


ChatGPT creates the three groups as:

<br>  - Late treated cohort: 35-40,
<br>  - Early treated cohort: 25-30,
<br>  - Pretreatment cohort: 20.

As observed, the ChatGPT groups do not match the age groups in the paper.

The first step to improve the ChatGPT prompt would be to explicitly mention the age range of the different categories.


**<br>2. There is redundancy with the regions dummy variables.**

ChatGPT correctly uses two out of the three regions dummies setting ´north´ as the reference. However, when running the regression accounting for fixed effects with the feols() function, the region dummies are included as explicit regressors. Normally, the regions dummies would need to account for fixed effect (on the right side of |). However, districts are already used to account for the regional fixed effect (by district). The addition of the regions dummy variable are therefore redundant and may cause bias estimations.

A second step we could take to improve th ecode would be to explicitely tell ChatGPT to only include the district fixed effect and not the region fixed effect.


**<br>3. Statistical significance does not account for the small sample size.**

Under Table 5 (p.24), it is indicated that "Statistical significance is evaluated relative to the small sample t distribution to account for the small number of clusters."

In its code ChatGPT did not account for the small sample size and simply applies the ´stars = TRUE´ function. This effect likely overstates the precision of the estimates.

A third step we could take would be to explicitely tell ChatGPT to account for the small sample size by evaluating statistical significance relative to the small sample t distribution in order to account for the small number of clusters.


**<br>4. The F-tests were not conducted by ChatGPT, which undermines the key point of the paper.**

A key part of table 4 is the F-tests conducted in order to assess which group had been more affected by migration $(\beta_2)$. By not conducting the F-test, ChatGPT undermines an important point of the paper that shows that the 1967 labor law change had a causal impact on the level of education of Malawi citizens.

A fourth step to take to improve ChatGPT's output would be to explicitly mention the necessity of the F-tests.


**<br>5. ChatGPT is very vague when producing the values of the coefficients and they do not math the actual code's output (hallucination).**

The coefficient estimations produced by ChatGPT are extremely vague. When we run the actual code in R we see that the coefficients are all underestimated and do not actually match the vague range produced by ChatGPT. It is likely that ChatGPT had not run the code it had produced and simply "copied and pasted" the values from Table 5 that we had included in our first prompt. This is likely an AI hallucination incident.

A fifth step to improve the prompt would be to tell ChatGPT to first run the code internally and then answer us.

<br>

**Improved prompt:**

Hey ChatGPT, I need your help replicating the results from Table 5 (p.24) in the attached paper. 

The age ranges for each group are: 

<br>  - Post treatment cohort: 20-24,
<br>  - Late treated cohort: 25-35,
<br>  - Early treated cohort: 35-44,
<br>  - Pretreatment cohort: 45-64.

When it comes to the fixed effects, please only include the district-level fixed effects and do not include the region-level fixed effects as these are redundant and could lead to estimation bias.

As you can see under table 5, it is written: "Statistical significance is evaluated relative to the small sample t distribution to account for the small number of clusters." I need you to account for this small sample size and to use the exact same aforementioned method. Do not take the general statistical significance.

Also, I need you to conduct F-tests between the different coefficients in order to evaluate whether the 1967 labor law had any impact on the level of educational attainment for Malawi citizens.

Now, take your time to think, I am not in a rush. Before you answer me with the relevant R code please run it internally and only report the values that you arrive to in your code.


Please use the codebook, dataset, and paper attached to this message to complete this task.

<br>

**b. Interpret the results. Differences with question 3.a.?**

Your answer.
